---
title: "Chaper 6 - Confidence Intervals"
author: "Daxiang Na"
date: "2022-10-06"
CJKmainfont: Songti SC
output: 
  pdf_document: 
    includes:
        header-includes:
                - \usepackage{xeCJK}
    latex_engine: xelatex
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70))
```

\

Q1: For sampling distribution, it can have sampling size $n$ and sampling time $m$, what determines if it follows CLT? $n$ or $m$? If we sample for 1 time and 100 times, each time with same size $n$, does that makes a difference?

# CLT and Sampling Distribution

```{r}
f <- function(x,N) {
  df <- replicate(N, rexp(x, 2))
  a <- df[,1]
  return(a)
}

a <- f(10,1)
b <- f(100,1)
c <- f(1000,1)
```

```{r}
set.seed(42)
library(dplyr)
f <- function(x) {
  N <- 30
  df <- as.data.frame(replicate(N, rexp(x, 2))) %>% 
    summarise_if(is.numeric, mean)
  a <- as.numeric(as.vector(df[1,]))
  return(a)
}
a <- f(10)
mean(a)
sd(a)
0.5/sqrt(10)

b <- f(20)
mean(b)
sd(b)
0.5/sqrt(20)

d <- f(30)
mean(d)
sd(d)
0.5/sqrt(30)

c <- rexp(100, 2)
mean(c)
sd(c)
0.5/sqrt(1)
```

# Margin of Error and Sample Size

$n=\left\lceil\frac{z_{\alpha / 2}^2 \cdot \sigma^2}{m^2}\right\rceil$

# Normal Distribution Curve with ggplot 正态分布曲线

```{r}
library(tidyverse)
f <- function(x, mu, sigma) {
return(1/(sqrt(2 *pi) * sigma) * exp(-0.5 * ((x - mu)/sigma)^2))
}

base <- ggplot() + xlim(-10, 10)
base + 
      geom_function(aes(colour = "mu = 0, sigma = 3"), fun = f, args = list(mu = 0, sigma = 3), size = 2, linetype = "longdash")
```

# t distribution curve with ggplot T 分布曲线

```{r}
base <- ggplot() + xlim(-10, 10)
base + 
      geom_function(aes(colour = "df = 5"), fun = dt, args = list(df=5)) +
      geom_function(aes(colour = "df = 20"), fun = dt, args = list(df=20)) +
      geom_function(aes(colour = "df = 500"), fun = dt, args = list(df=500)) 
```
