---
title: "Chapter 12 Non-parametric Test"
author: "Daxiang Na"
date: "2022-10-25"
CJKmainfont: STKaiti
output: 
  pdf_document: 
    includes:
        header-includes:
                - \usepackage{xeCJK}
    latex_engine: xelatex
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70))
library(tidyverse)
library(knitr)
```

# Questions:

## What is $V$ in the returned result of `wilcox.test`

```{r, echo = T}
##
  ##  Wilcoxon signed rank exact test
  ##
## data: data$air and data$sulf.diox
## V = 21, p-value = 0.006653
## alternative hypothesis: true location shift is not equal to 0
```

## What is the definition of T in `psignrank`?

In the `psignrank` document(<https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SignRank>), the the Wilcoxon signed rank statistic is " the sum of the ranks of the absolute values x[i] for which x[i] is positive". I am wondering if $T = T^+$ instead of $T = min(T^+, T^-)$? This appears to be correct when solving the example problem, where `2*(1 - psignrank(75.5,n=14))` = 0.135 but `2*psignrank(29.5,n=14)` != 0.135

I therefore wonder if $T = T^+$ should be the definition of `psignrank` in R.

[Another question regarding this topic]{.underline}: when testing the two-tailed hypothesis, when should we use `2*(1 - psignrank(T,n))` and `2*psignrank(T, n)`? i.e. what is the mean for the wilcoxon signed rank distribution? Is it $n(n + 1)/4$ as stated in the document?

**This is right.**

So when we calculate the p value, we determine whether to use `2*(1 - psignrank(T,n))` or `2*psignrank(T, n)` based on if $T > n(n + 1)/4$ or $T < n(n + 1)/4$, right?

## How can we decide which probability to calculate in one-sided Wilcoxon Rank Sum test?

In the Wilcoxon Rank Sum test, we always use $W = min(W_1, W_2)$ and $\mu_W = n_{min}*(n_{min} + n_{large} + 1)/2$ no matter if we want to know if $H_1: \mu_1 - \mu_2 < 0$ or $H_1: \mu_2 - \mu_1 < 0$, how to we determine if p-value = $1 - Pr(z < z_W)$ or p-value = $Pr(z < z_W)$?

My understanding: it depends on which group has $W_{min}$. In the Wilcoxon Rank Sum test we are calculating the probability of getting $W_{min}$ in the distribution of $N(n_{min}*(n_{min} + n_{large} + 1)/2, sigma)$. In the example problem in the slides, H1 is about group 1 has less values, and group 1 turns out to have $W_{min}$, therefore p-value = Pr(z < 0). 

**This understanding is correct**. The p-value will be determined by which group as $W_{min}$ and what $H_1$ is about.

## Question about CLT

For sampling distribution, it can have sampling size $n$ and sampling time $m$, what determines if it follows CLT? $n$ or $m$? If we sample for 1 time and 100 times, each time with same size $n$, does that makes a difference? Another way to ask this question: if we sample for 3 times ($m = 3$), but each time with sample size $n = 1000$, does that follow CLT?

**Answer**: n determines the distribution of sampling. $\bar x \sim N(\mu, \sigma/\sqrt n)$. The change of m does not affect the distribution. What m makes a difference: when m = 1, you only see one $\bar x$ on histogram; when m = 100, you will see 100 $\bar x$ on histogram, and their distribution will be closer to the bell shape than m = 1. However, no matter if m = 1 or m = 100, their distribution does not change.

## is it okay to share the Infererence cheat sheet raw file?

Willing to extend it.

\

# Wilcoxon Signed Rank Test

-   Only for paired sample.

-   Evaluate the null hypothesis: $Z_{T} = (T - \mu_{T})/\sigma_{T}$

-   Note: $$
    \begin{aligned}
    &\mu_T=0 \\
    &\sigma_T=\sqrt{\frac{n(n+1)(2 n+1)}{6}}
    \end{aligned}
    $$

-   When n is large enough ($n > 12$), we get$$
    Z_T \sim N(0,1)
    $$

-   calculate the probability of getting $Z_{T}$ when $\mu = 0$ is true.

-   For two-sided test, follow what we do in the sampling distribution:

    -   2\*p when z \< 0

    -   2\*(1 - p) when z \> 0

-   if $n > 12$ , you can just apply CLT, the R code is: `wilcox.test(before, after, paired = T, exact = F, correct = F)`. `exact =` determines if the statistics follow normal distribution (`exact = F`) or exact distribution (`exact = T`).

-   If $n \le 12$, we cannot use the normal approximation. In that case, we use `psignrank(T,n)` in R to calculate the exact distribution.

    -   \textcolor{red}{R requires $T = T^+$ for this to work correctly!}

# Wilcoxon Rank-Sum test (also known as Mann-Whitney U test)

-   nonparametric analog to the two-sample t-test

-   get $W_{1}$ and $W_{2}$

-   $W = min(W_1, W_2)$

-   $n_1$ = sample size with the [smaller]{.underline} sum of ranks.

-   $n_2$ = sample size with the [larger]{.underline} sum of ranks.

-   $$
    \mu_W=\frac{n_1\left(n_1+n_2+1\right)}{2} \text { and } \sigma_W=\sqrt{\frac{n_1 n_2\left(n_1+n_2+1\right)}{12}}
    $$

-   $$
    z_W=\frac{W-\mu_W}{\sigma_W}
    $$

-   $z_W \sim N(0, 1)$ when $n_1$ and $n_2$ are large enough (n1, n2 \> 10).

    -   in R: `wilcox.test(..., exact = F, correct = F, paired = F, alt = "")`

-   When $n_1$ and $n_2$ are very small (i.e. [either is less than or equal to 10]{.underline}), we can use the exact distribution to calculate the p-values. In R: `pwilcox(Wobs, n1, n2)`

    -   in this case, $W_{obs} = W - n_1(n_1 + 1)/2$
    -   `wilcox.test` also works when `exact = T`

-   correct: correct the data with continuity correction
